\section{Introduction}
\label{sec:intro}

Simultaneous speech translation (SST) must commit target-language output before
hearing the future of the source stream. Every design choice in the field ---
wait-$k$ schedules \citep{ma2019stacl}, adaptive READ/WRITE policies
\citep{liu2020lowlatency,papi2023alignatt}, long-form streaming architectures
\citep{ouyang2025infinisst} --- is a way of coping with this single constraint:
at commit time, the model knows only a prefix. The words most damaged by this
constraint are the ones that matter most in talks and lectures: rare
terminology, named entities, numbers --- exactly the content a translator cannot
guess from a truncated prefix and cannot revise after committing.

Yet in the settings where SST is most needed --- conference talks, lectures,
screen-shared meetings --- the future is, quite literally, projected on the
wall. Speakers put up a slide and then discuss it, typically for 30--60 seconds;
professional simultaneous interpreters exploit this routinely, requesting decks
in advance and reading the projected slides from the booth. The visual channel
is thus a \emph{legal lookahead}: it reveals upcoming terminology before it is
spoken, at zero cost to the audio decision loop, because slides change on a
timescale of tens of seconds and can be processed asynchronously off the
critical path. We call this \emph{vision-as-anticipation}, and this paper builds
the method, the evidence, and the evaluation protocol for it.

Prior work has approached the ingredients separately but never the combination.
Slide context improves \emph{offline ASR} of conference talks
\citep{sinhamahapatra2025slides}, but transcription has no latency dimension and
cannot exploit the cross-lingual structure of slides. Images help
\emph{simultaneous text MT} on caption datasets \citep{caglayan2020simultaneous,
ive2021exploiting}, establishing the anticipation insight in a toy regime ---
single photographs, no speech, sentence-level decoding. Most recently,
OmniFusion \citep{koneru2025omnifusion} fuses a multimodal foundation model with
a translation LLM for En$\rightarrow$X SimulST, but processes the image
\emph{synchronously on the decoding path} (paying added latency for vision) and
attributes its speed-up to removing the ASR--MT cascade, not to the visual
channel itself. No existing system treats vision as a zero-latency anticipation
channel for speech translation, and no existing evaluation separates
vision-induced gains from architectural ones.

We decompose the benefit of slide evidence into three mechanisms:
\textbf{M1 anticipation} (slides temporally precede the speech that discusses
them; direction-agnostic), \textbf{M2 recognition support} (source-language
slide terms help the model resolve rare or mis-heard words), and
\textbf{M3 target-form supply} (when the slide language equals the target
language --- common in X$\rightarrow$En, since English is the lingua franca of
slides --- the slide hands the system the target surface form of an upcoming
term, turning translation into recognition plus copying). The decomposition is
testable: slide language is a per-talk stratification variable, and M3 predicts
larger terminology gains and higher slide-string copy rates precisely on talks
whose slides are in the target language.

A controlled probe validates the premise before any system building. On
Greek$\rightarrow$English TEDx segments (English slides, human references) and
Chinese lecture segments with a Local Agreement streaming policy
\citep{liu2020lowlatency}, injecting the reference's own rare terms --- a
perfect-anticipation oracle --- improves chrF by $+7.0$ pooled ($p{=}0.03$) and
$+18.8$ on segments where the baseline fails on terminology ($p{<}10^{-4}$;
term recall $0.12\!\rightarrow\!0.41$). The realistic channel initially
delivers none of this: OCR-extracted slide terms are neutral-to-harmful. But
replacing OCR with a VLM slide reader recovers $+14.1$ chrF on hard segments
(75\% of the oracle gain), and adding need-based injection --- only on segments
where the baseline is predicted to fail --- recovers $+7.3$ chrF pooled
($p{<}0.001$), statistically indistinguishable from the oracle upper bound.
The chain isolates the two problems a method must solve: \emph{what} to extract
(VLM, not OCR) and \emph{when} to inject (a need predictor; we show a naive
LLM relevance gate has no selectivity), while a wrong-slide control shows
prompt-level bad evidence is benign for strong models.

Our contributions:
\begin{enumerate}[leftmargin=*]
    \item \textbf{Setting and mechanism analysis.} We formulate vision-as-anticipation
    SST --- an asynchronous slide channel off the audio critical path --- and
    decompose its benefit into anticipation, recognition support, and
    target-form supply, each separately measurable via slide-language
    stratification (\S\ref{sec:task}).
    \item \textbf{Benchmark suite.} A direction-general evaluation over real
    slides only: mTEDx-V (long-form X$\rightarrow$En, live video-recoverable
    talks, human references, visual-signal stratification), Chinese-LiPS-Long
    (dedicated 1080p slide stream rebuilt on the original session timeline),
    and ACL 60/60 (En$\rightarrow$X control with third-party tagged
    terminology), with a Wikidata-based cross-lingual term protocol
    (\S\ref{sec:benchmark}).
    \item \textbf{Method and evidence.} An asynchronous slide-evidence pipeline
    (change-triggered VLM reading, term filtering, need-gated injection) whose
    probe-scale ablation closes the full oracle gap, and an analysis showing
    that need prediction --- not extraction --- is the remaining open problem
    (\S\ref{sec:method}, \S\ref{sec:results}).
\end{enumerate}
