\section{Experimental Setup}
\label{sec:experiments}

\subsection{Probe protocol (\S\ref{sec:probe-results})}

The controlled probe isolates the information value of slide evidence from
system engineering. Streaming is simulated over incrementally revealed source
transcripts with the \textbf{Local Agreement} policy \citep{liu2020lowlatency}:
at each READ step (2 words for Greek; 4 characters for Chinese) the translator
produces a full translation of the prefix, and the longest common prefix of two
consecutive outputs beyond already-committed words is committed (monotonic, no
retraction; final step appends the tail). The translator is
\textbf{Qwen3-32B} (temperature 0, thinking disabled) served with vLLM on one
H200; the pilot replication uses Qwen2.5-7B-q4. Text-prefix simulation
deliberately removes ASR noise so that evidence effects are attributable to
translation, not recognition; the full-system evaluation adds the audio front
end.

\paragraph{Data.} 40 segments of a Greek TEDx talk with English slides
(mTEDx-V; M3 regime; human references from TED translators) and 16 Chinese
lecture segments (Chinese-LiPS; M1+M2 regime; author-drafted references),
segment length $\geq 8$ reference words, sampled past the first minute.

\paragraph{Conditions.} (A) audio-only baseline; (B) slide terms from the
currently visible slide, extracted by 480p tesseract OCR or by a VLM reader
(Qwen2.5-VL-7B) over the same frames with a 90\,s lookback; (C) oracle future
terms: the reference's own rare content words --- a perfect-anticipation upper
bound; (D) wrong-slide terms from a different moment of the talk (faithfulness
control). Gating variants inject B's terms selectively: an oracle need gate
(hard stratum only), a learned need-predictor (below), an LLM relevance gate,
and rule-based uncertainty gates. A decode-level variant (\emph{soft
injection}) applies the slide terms as a logit bias on hint tokens instead of
prompt text, leaving the prompt untouched.

\paragraph{Metrics.} chrF against references (sacreBLEU); term recall over the
reference's rare terms; Average Lagging \citep{ma2019stacl} in READ units.
Significance: paired bootstrap over segments ($10^4$ resamples, one-sided).
The \emph{hard stratum} (baseline term recall $\leq 0.4$) is fixed before any
condition comparison.

\subsection{Learned need-predictor}

Failure labels are mined automatically: each of 1{,}540 Greek segments (18
talks, test+valid) is translated offline by the same model; a segment is
\emph{hard} if term recall $\leq 0.4$ (24\% base rate). Features must be
online-computable: source-side surface statistics (word-length profile,
long-word rates, digits, character diversity) and model-confidence features
from the draft's token logprobs (mean, min, low-confidence fractions).
A class-balanced logistic regression is trained with \textbf{talk-held-out}
evaluation (the probe talk is never seen in training); the decision threshold
is calibrated on training talks. End-task evaluation plugs the predictor's
per-segment decisions into the gated condition on the probe.

\subsection{Full-system evaluation (planned)}

The complete pipeline (streaming audio front end, asynchronous slide worker,
gated/soft injection) is evaluated on the three benchmark strata of
\S\ref{sec:benchmark}: S1 mTEDx-V (X$\rightarrow$En long-form, stratified by
VLM-derived visual-signal class and slide language), S2 Chinese-LiPS-Long
(original-timeline sessions, dedicated 1080p slide stream), S3 ACL 60/60
(En$\rightarrow$Zh control with third-party tagged terms). Quality
(chrF/XCOMET) is reported jointly with computation-aware Average Lagging;
terminology follows the Wikidata cross-lingual protocol with per-stratum
coverage; faithfulness metrics are wrong-slide adoption rate and
visible-but-unspoken rate.
\todo{Finalize full-system decoding parameters and hardware/cost details.}
