\section{Results}
\label{sec:results}

\subsection{Probe: the anticipation-value chain}
\label{sec:probe-results}

Table~\ref{tab:probe} reports the controlled probe (\S\ref{sec:experiments}) on
Greek$\rightarrow$English TEDx segments with English slides (M3 regime, human
references, $n{=}40$; Qwen3-32B, Local Agreement policy). The \emph{hard
stratum} contains segments where the audio-only baseline misses most reference
terminology (term recall $\leq 0.4$; $n{=}22$; the hard rate, $\sim$24\%, is
stable across 32B and 14B models --- a data property).

\begin{table}[t]
\centering
\small
\begin{tabular}{lcccc}
\toprule
Condition & chrF & $\Delta$ pooled & $p$ & $\Delta$ hard \\
\midrule
A\; audio-only                & 38.8 & ---    & ---   & ---     \\
B\; slide terms (OCR)         & 35.2 & $-2.9$ & n.s.  & $+2.3$  \\
B\; slide terms (VLM)         & 39.9 & $-0.6$ & n.s.  & $+14.1$ \\
B\; VLM + need-gate (oracle)  & \textbf{45.3} & $\mathbf{+7.3}$ & $<.001$ & $+13.3$ \\
C\; oracle future terms       & 46.6 & $+7.0$ & .059  & $+16.3$ \\
D\; wrong-slide terms         & 38.6 & $-0.2$ & n.s.  & ---     \\
\bottomrule
\end{tabular}
\caption{Anticipation-value probe (el$\rightarrow$en, $n{=}40$; paired
bootstrap vs.\ A; vLLM serving). Slide evidence read by a VLM and injected
selectively matches the oracle-terms upper bound.}
\label{tab:probe}
\end{table}

\paragraph{The premise holds, and gains sit exactly where predicted.}
Injecting the reference's own rare terms (C) --- a perfect-anticipation oracle
--- yields $+7.0$ chrF pooled and $+18.8$ on the hard stratum ($p{<}10^{-4}$),
with hard-stratum term recall rising $0.12\!\rightarrow\!0.41$ ($3.4\times$).
Gains concentrate on segments where the baseline fails on terminology and are
neutral elsewhere. The effect replicates across model scales (7B, 14B, 32B)
and across all three benchmark regimes (below).

\paragraph{Extraction quality is load-bearing.} OCR slide terms are
neutral-to-harmful (480p fragments); VLM-read terms recover $+14.1$ chrF on
hard segments --- 75\% of the oracle's hard-stratum gain from the \emph{real}
visual channel.

\subsection{Injection form: a three-step reversal}
\label{sec:injection-form}

Our initial runs suggested prompt hints disturb easy segments (motivating
need-gates and decode-level soft injection); the investigation closed in a
reversal that reshapes the method:

\begin{enumerate}[leftmargin=*]
\item \textbf{Apparent easy-segment harm} under vLLM serving affected
  \emph{all} hint conditions (including wrong-slide) with truncation
  signatures, and an oracle need-gate recovered $+7.3$ pooled ($p{<}.001$).
\item \textbf{Decode-level soft injection is benign but powerless}: uniform
  logit bias saturates at $+1.3$ hard ($+8$ bias), and trie-constrained
  shallow fusion --- boosting a term's continuation only along its prefix ---
  reaches only $+0.7$ pooled (termR $0.29$ vs.\ prompt's $0.46$).
  Terminology failures are \emph{phrase-planning} failures, not single-token
  near-misses; posterior logit nudges are too late and too local, while prompt
  hints act at the planning level.
\item \textbf{The root cause was serving nondeterminism}: batched vLLM serving
  returns different outputs for identical prefixes, stalling the
  longest-common-prefix commit. Under deterministic decoding the baseline
  jumps ($38.8\!\rightarrow\!62.7$ corpus chrF on the same segments) and
  prompt injection is simultaneously \emph{powerful and benign}: on 50
  14B-hard segments, $+6.5$ chrF ($p{<}0.001$), term recall
  $0.27\!\rightarrow\!0.46$, easy-$\Delta$ $\approx 0$.
\end{enumerate}

\noindent\textbf{Full-system verdict.} Deterministic (or drift-tolerant)
commit + prompt-level hint injection; need-gating demotes from a necessity to
a token-budget optimization; trie biasing is recorded as an informative
negative result. Practically: simultaneous pipelines built on serving-grade
inference must either pin deterministic kernels or use drift-tolerant commit
policies --- a deployment finding independent of the visual channel.

\subsection{Cross-regime mechanism evidence (H2)}
\label{sec:mechanism}

The En$\rightarrow$Zh control stratum (ACL 60/60, 5 talks, 60 segments,
official tagged terminology; slides are \emph{source}-language $\Rightarrow$
M1+M2 only) delivers the first direct test of the mechanism decomposition:

\begin{table}[t]
\centering
\small
\begin{tabular}{lcccc}
\toprule
Condition (En$\rightarrow$Zh) & chrF & $\Delta$ & termAcc & $\Delta$ hard \\
\midrule
none              & 26.3 & ---    & 0.42 & ---    \\
\textbf{slide (VLM, ungated)} & \textbf{28.3} & $+3.1$ & 0.44 & $+9.1$ \\
oracle terms      & 25.9 & $-1.1$ & 0.39 & $+3.8$ \\
wrong slide       & 23.1 & $-3.2$ & 0.32 & $+3.3$ \\
\bottomrule
\end{tabular}
\caption{S3 control (ACL 60/60, $n{=}60$, official tagged terms; vLLM-era
protocol, zh char-level LCP --- truncation caveat applies).}
\label{tab:s3}
\end{table}

\begin{itemize}[leftmargin=*]
\item \textbf{Copy-rate 0.00} (2/462): English slide strings never enter the
  Chinese output verbatim, versus the active copy channel in X$\rightarrow$En
  --- exactly H2's prediction that M3 (target-form supply) exists only when
  slide language equals the target language, while M2 (recognition support)
  gains persist ($+9.1$ hard) without copying.
\item \textbf{Real slides transfer ungated} on this stratum: ACL screen-recorded
  slides are dense (95\% of frames term-dense, $8\times$ mTEDx) and
  speech-aligned (speakers read their slides), so injection matches the
  model's natural phrasing.
\item \textbf{Faithfulness is direction-sensitive}: wrong-slide evidence is
  benign for 32B X$\rightarrow$En but harmful En$\rightarrow$Zh ($-3.2$);
  faithfulness metrics must be reported per direction.
\end{itemize}

\subsection{Learned need prediction (as budget optimization)}
Surface features of the source predict terminology failure at chance
(talk-held-out AUC 0.49--0.52); draft-translation logprob features reach AUC
0.62 --- the model's own uncertainty carries the signal --- preserving the full
hard-stratum gain end-to-end with imperfect precision (0.16). Under the
deterministic-decoding verdict this predictor is no longer required for
safety, but remains useful to prioritize hint budget in long sessions.

\subsection{Full-system results}
\todo{Main table over S1/S2/S3 with the deterministic streaming pipeline:
quality $\times$ computation-aware latency, Wikidata term protocol with
coverage, copy-rate by slide-language regime, per-direction faithfulness.}
